---
title: "Satnogs Analysis"
author: "Ada Lazuli"
date: '2022-10-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r}
library(ggplot2)
library(stringr)
library(dplyr)
```


# Background

The Satellite Network Open Ground Stations (SATNOGS), was created in 2014 as part of a NASA Space App Challenge. Since then, the project has grown to over 200 sites around the world.

The project targets satellites in a Low Earth Orbit (LEO) that transmit in the Very High Frequency (VHF) and Ultra High Frequency (UHF) ranges.  

The SATNOGS project collects the observations from the ground stations and makes low resolution power spectral displays (PSDs) rendering of the signals available. 

Since 2014, the SATNOGS network has collected over 6 million signals and presents itself as a wonderful source of labeled data for machine learning applications.

A [SATNOGS Webscraper](https://github.com/Maple-Lazuli/satnogs-webscraper-v2) was used to pulls around 1 million observations from SATNOGS

# Data Loading

The data was uploaded to GitHub in small chunks and will need to be re-combined. This was accomplished using `cat` to concatenate the scrapped CSVs into __bad.csv__ and __good.csv__, where __bad.csv__ contains observations that lack signals and __good.csv__ contains observations that contain signals.

```{bash, engine.opts='-l'}
rm ../data/bad.csv
rm ../data/good.csv

cat ../data/bad.csv.part* >> ../data/bad.csv
cat ../data/good.csv.part* >> ../data/good.csv
```

Next the data was loaded into R data frames, with a feature being added that indicates whether or not a signal is present.

```{r}
bad <- read.csv("../data/bad.csv")
bad$Signal <- FALSE
good <- read.csv("../data/good.csv")
good$Signal <- TRUE
df <- rbind(bad, good)
```

# Data Description

The data consists of a total of `r nrow(df)` observations with the following features for each:

1. Observation ID - The unique numeric ID that SATNOGS assigned the observation event.
2. Satellite - The name of the satellite from which the signals originated, having the form '<SSC #> - Common Name'.
3. Station - The ground station that intercepted the signal.
4. Status - The overall status of the observation, such as good, bad, failed, unknown.
5. Status Message - The numeric representation of the status.
6. Transmitter - Nuanced details about the signal used for transmission. 
7. Frequency - The frequency of the signal in Hertz (Hz).
8. Mode - High level detail on the signal used for transmission.
9. Metadata - Free Form field encoded as a JSON that captures details of the observation.
10. Downloads - Downloadable content, such as Audio, Waterfalls, or Data
11. Waterfall_Status - The vetted status of the waterfall, such as signal present.
12. Signal - Boolean value indicating whether a signal is present in the waterfall.

```{r}
glimpse(df)
```


## Frequency Description

SATNOGS focuses on signals transmitted in VHF and UHF, however it does contain observations of signals that were transmitted in L-Band and S-Band. 

To better understand the frequency coverage provided by SATNOGS, the data needed to be grouped into bins for each band, with the following criteria used for labeling:

1. __VHF__ if the signal was between 30 and 300 MHz
2. __UHF__ if the signal was between 300 and 100 MHz
3. __L-Band__ if the signal was between 1000 and 2000 MHz
4. __S-Band__ if the signal was between 2000 and 4000 MHz
5. __UNK__ if the signal was in a range not specified previously



```{r}
label_band <- function(freq) {
if ((freq <= 300) & (freq >= 30))
return("VHF")
if ((freq <= 1000 ) & (freq > 300))
return("UHF")
if ((freq <= 2000) & (freq > 1000))
return("L-Band")
if ((freq <= 4000) & (freq > 2000))
return("S-Band")
return("Unk")
}
df <- df %>% mutate(Frequency = as.numeric(gsub(r"[\D]", "", Frequency))) %>%
  mutate(Frequency = Frequency/1000000)
df$Band <- sapply(df$Frequency, FUN = label_band)
counts <- df %>% group_by(Signal, Band) %>% count()
ggplot(counts, aes(x = Band, y = n, color = Signal, fill = Signal)) + geom_col(position = "dodge", alpha = 0.5) + labs(title = "Frequency Bands", x ="Band", y = "Counts") + coord_flip()
ggplot(counts, aes(x = Band, y = n, color = Signal, fill = Signal)) + geom_col(position = "dodge", alpha = 0.5) + labs(title = "Frequency Bands", x ="Band", y = "Logarithmic Counts") + coord_flip() + scale_y_log10()
```



### VHF

```{r}
ggplot(df %>% filter(Band == "VHF"), aes(x = Frequency)) + geom_histogram(bins = 1000) +  labs(title = "Histogram of VHF Frequencies", x ="Frequency", y = "Counts")
ggplot(df %>% filter(Band == "VHF"), aes(x = Frequency, color = Signal, fill = Signal)) + geom_histogram(bins = 1000, , alpha = 0.3) +  labs(title = "Histogram of VHF Frequencies", x ="Frequency", y = "Counts")
```

### UHF

```{r}
ggplot(df %>% filter(Band == "UHF"), aes(x = Frequency)) + geom_histogram(bins = 1000) +  labs(title = "Histogram of UHF Frequencies", x ="Frequency", y = "Counts")
ggplot(df %>% filter(Band == "UHF"), aes(x = Frequency, color = Signal, fill = Signal)) + geom_histogram(bins = 1000, alpha = 0.4) +  labs(title = "Histogram of UHF Frequencies", x ="Frequency", y = "Counts")
```


### L-Band

```{r}
ggplot(df %>% filter(Band == "L-Band"), aes(x = Frequency)) + geom_histogram(bins = 1000) +  labs(title = "Histogram of L-Band Frequencies", x ="Frequency", y = "Counts")
ggplot(df %>% filter(Band == "L-Band"), aes(x = Frequency, color = Signal, fill = Signal)) + geom_histogram(bins = 1000, alpha = 0.3) +  labs(title = "Histogram of L-Band Frequencies", x ="Frequency", y = "Counts")
```


### S-Band

```{r}
ggplot(df %>% filter(Band == "S-Band"), aes(x = Frequency)) + geom_histogram(bins = 1000) +  labs(title = "Histogram of S-Band Frequencies", x ="Frequency", y = "Counts")
ggplot(df %>% filter(Band == "S-Band"), aes(x = Frequency, color = Signal, fill = Signal)) + geom_histogram(bins = 1000, alpha = 0.3) +  labs(title = "Histogram of S-Band Frequencies", x ="Frequency", y = "Counts")
```


### Unk

```{r}
ggplot(df %>% filter(Band == "Unk"), aes(x = Frequency)) + geom_histogram(bins = 1000) +  labs(title = "Histogram of Un-Labeled Frequency Bands", x ="Frequency", y = "Counts")
ggplot(df %>% filter(Band == "Unk"), aes(x = Frequency, color = Signal, fill = Signal)) + geom_histogram(bins = 1000, alpha = 0.3) +  labs(title = "Histogram of Un-Labeled Frequency Bands", x ="Frequency", y = "Counts")
```


## Modulation Types

```{r}
df <- df %>% mutate(Mode = str_sub(Mode, 2, -2)) %>%  mutate(Mode = str_extract(Mode, "[A-Z]+"))
counts <- df %>% group_by(Signal, Mode) %>% count()
ggplot(counts, aes(x = Mode, y = n, color = Signal, fill = Signal)) + geom_col(position = "dodge", alpha = 0.5) + labs(title = "Modulation Performance", x ="Modulation Type", y = "Counts") + coord_flip()
ggplot(counts, aes(x = Mode, y = n, color = Signal, fill = Signal)) + geom_col(position = "dodge", alpha = 0.5) + labs(title = "Modulation Performance", x ="Modulation Type", y = "Logarithmic Counts") + coord_flip() + scale_y_log10()

```

## Satellites

```{r}
counts <- df %>% group_by(Satellite) %>% count()
```

There are exactly `r length(unique(df$Satellite))` unique satellites, with the 5-Number summary for the observation count of each satellite is presented below:

```{r}
summary(counts$n)
ggplot(counts, aes(x = n)) + geom_histogram(bins = 1000) +  labs(title = "Histogram of Satellite Observation Occurences", x ="Number of Observations", y = "Occurences")
```

